PyPy's Approach to Virtual Machine Construction

### 2006-07-12 talk/dls2006/paper.tex
tweaking the RPython feature enumeration

There are two major components in PyPy:
%
\begin{enumerate}
\item the \textit{Standard Interpreter}: an implementation of the Python programming
language, mostly complete and compliant with the current version of the
language, Python 2.4.
\item the \textit{Translation Process}: a translation tool-suite whose goal is to
compile subsets of Python to various environments.
\end{enumerate}
%
In particular, we have defined a subset of the Python language called
``restricted Python'' or RPython.  This sublanguage is not restricted
syntactically, but only in the way it manipulates objects of different
types.  The restrictions are a compromise between the expressivity and
the need to statically infer enough type information to generate
efficient code.  RPython still supports inheritance but limited to
single inheritance with some mix-in support, dynamic dispatch, to some
extent keywords arguments and varargs, first-class function and class
values, limited use of bound methods, runtime \texttt{isinstance} and type
queries, but no runtime reflection, further bindings in class and
global namespaces are assumed constant. RPython code can be run on a
Python interpreter without severe semantics mismatches, figure
\ref{fig_create_frame} shows some RPython code from the
\textit{Standard Interpreter}, it is still quite idiomatic Python code
using the built-in dictionary type and first-class class values
instantiation.

>>>>

There are two major components in PyPy:
%
\begin{enumerate}
\item the \textit{Standard Interpreter}: an implementation of the Python programming
language, mostly complete and compliant with the current version of the
language, Python 2.4.
\item the \textit{Translation Process}: a translation tool-suite whose goal is to
compile subsets of Python to various environments.
\end{enumerate}
%
In particular, we have defined a subset of the Python language called
``restricted Python'' or RPython.  This sublanguage is not restricted
syntactically, but only in the way it manipulates objects of different
types.  The restrictions are a compromise between the expressivity and
the need to statically infer enough type information to generate
efficient code.  RPython still supports exceptions, inheritance but
limited to single inheritance with some mix-in support, dynamic
dispatch, to some extent keywords arguments and varargs, first-class
function and class values, limited use of bound methods, runtime
\texttt{isinstance} and type queries, but no runtime
reflection. Bindings in class and global namespaces are assumed
constant. RPython code can be run on a Python interpreter without
severe semantics mismatches, figure \ref{fig_create_frame} shows some
RPython code from the \textit{Standard Interpreter}, it is still quite
idiomatic Python code using the built-in dictionary type and
first-class class values instantiation.
